### 2021-06-22 article/article.tex
v0.4.0. Sweep parameter now array, initital states

\section{Abstract}
	\texttt{spinsim} is a \emph{python} package that simulates spin half and spin one quantum mechanical systems following a time dependent Shroedinger equation. It makes use of \texttt{numba.cuda} \cite{lam_numba_2015}, which is an \emph{LLVM} (Low Level Virtual Machine) \cite{lattner_llvm_2004} compiler, and other optimisations, to allow for fast and accurate evaluation on \emph{Nvidia Cuda} \cite{nickolls_scalable_2008} compatible systems using GPU parallelisation. \texttt{spinsim} is available for installation on \emph{PyPI}, and the source code is available on \emph{github}. The initial use case for the package will be to simulate quantum sensing-based Bose Einstein condensate (BEC) experiments for the Monash University School of Physics and Astronomy spinor BEC lab, but we anticipate it will be useful in simulating any range of spin half or spin one quantum systems with time dependent Hamiltonians that cannot be solved analytically. These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin one systems of nitrogen vacancy centres (NVCs) and BECs.

@@

\section{Implementation and architecture}
	% \subsection{Mathematical methods}
	\subsection{Quantum mechanics background}
		\texttt{spinsim} solves the time-dependent Schr\"{o}dinger equation

		\begin{align}
			i\hbar\frac{\mathrm{d}\psi(t)}{\mathrm{d}t} &= H(t)\psi(t),\label{eq:schroedinger}
		\end{align}

		where the quantum state \(\psi(t) \in \mathbb{C}^N\) is assumed normalised, and the Hamiltonian \(H(t) \in \mathbb{C}^{N \times N}\) is Hermitian. Here \(N\) is the number of levels in the quantum system. Often we are considering systems with spin \(S\). Here spin-half is the \(N = 2\) case, and spin-one is \(N = 3\). We set \(\hbar = 1\), so that the Hamiltonian has units

		% Furthermore, we have chosen for the Hamiltonian \(H(t)\) to have the physical dimension of frequency, rather than the more typical dimension of energy. To elaborate, we represent energies \(E\) of the system as their equivalent frequencies \(f\) from the Planck relation

		\begin{align}
			E &= \hbar \omega\\
			&= 2\pi\hbar f,
		\end{align}

		% where \(\hbar\) is (reduced) Planck's constant (or equivalently, setting 
		% \(\hbar = 1\)). In addition, rather than being represented by standard coordinates in \(\mathbb{C}^{N \times N}\), in \texttt{spinsim} the Hamiltonian \(H(t)\) is instead represented with respect to a choice of basis for the corresponding Lie Algebra, \(\mathfrak{su}(N)\). For example, when set to spin half mode, the \texttt{spinsim} package solves the time dependent Schroedinger equation of the form
		Rather than ... parametrizing the problem ... we consider time varying real coefficients in a linear combination of fixed operators,

		\begin{align}
			H(t) &= \sum_{j = 1}^{N^2 - 1} a_j(t) A_j.
		\end{align}

		It is well known that ... spin in magnetic field has ... . We exclude the identity ... energy zero point.

		In the spin one case there is no standard basis ... . Choices include Gell Mann, quadrupole basis, ... . In general ... Lie algebra ... mathfrak ... admits a basis ... .
		% \begin{align}
		% 	\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z) \psi(t),
		% \end{align}

		where \(i^2 = -1\), \(\psi(t) \in \mathbb{C}^2\), and the spin half spin projection operators are given by

		DONT WRITE OUT, sigma
		\begin{align}
			J_x &= \frac12\begin{pmatrix}
				0 & 1 \\
				1 & 0
			\end{pmatrix},
			&J_y &= \frac12\begin{pmatrix}
				0 & -i \\
				i &  0
			\end{pmatrix},
			&\textrm{and }J_z &= \frac12\begin{pmatrix}
				1 &  0 \\
				0 & -1
			\end{pmatrix}.\label{eq:spin_half_operators}
		\end{align}

		Call omega, no implementation detail
		% The fields \(f\) that represents the time dependent Hamiltonian \(H(t)\), are the collection of energy functions \(f_x(t), f_y(t), f_z(t)\), with time \(t\) and \(f\) with physical dimension of frequency that control the dynamics of the system. The user must define a method that returns a sample of these field functions when a sampling time is input. In physical terms, these functions could represent the \(x,y,z\) components of a magnetic field applied to a magnetically spin one BEC or NVC.

		Similarly, when \texttt{spinsim} is set to spin one mode, it can solve the Schroedinger equation of the form

		\begin{align}
			\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z + f_q(t) Q) \psi(t).
		\end{align}

		where now \(\psi(t) \in \mathbb{C}^3\), and the spin one operators are given by

		\begin{align}
			J_x &= \frac{1}{\sqrt{2}}\begin{pmatrix}
				0 & 1 & 0 \\
				1 & 0 & 1 \\
				0 & 1 & 0
			\end{pmatrix},&
			J_y &= \frac{1}{\sqrt{2}}\begin{pmatrix}
				0 & -i &  0 \\
				i &  0 & -i \\
				0 &  i &  0
			\end{pmatrix},\nonumber\\
			J_z &= \begin{pmatrix}
				1 & 0 &  0 \\
				0 & 0 &  0 \\
				0 & 0 & -1
			\end{pmatrix},&
			\textrm{and }Q &= \frac{1}{3}\begin{pmatrix}
				1 &  0 & 0 \\
				0 & -2 & 0 \\
				0 &  0 & 1
			\end{pmatrix}.\label{eq:spin_one_operators}
		\end{align}

		The matrices \(J_x, J_y, J_z\) are regular spin operators, and \(Q\) is a quadrupole operator. Note that \(Q\) is proportional to \(Q_{zz}\) as defined by Hamley et al \cite{hamley_spin-nematic_2012}, and \(Q_0\) as defined by Di et al \cite{di_dipolequadrupole_2010}.
		INLINE Q DIAG. WHY WE'VE PICKED THESE FOUR. For ground state spin one atoms in a magnetic field, ... Breit Rabi ... . COULD BE READILY EXTENDED TO INCLUDE FURTHER ELEMENTS OR TO HIGHER SPINS.

		% Too far out of scope
		% Alternative: integrate Bloch equations 3 coupled real equations
		% fast and parallel function from wavefunctions
		Frequently when one calculates the times series of the quantum state \(\psi(t)\), they are also interested in its expected spin projection \(\left\langle J\right\rangle\)(t). It is given by the vector

		% Inline
		\begin{align}
			\left\langle J\right\rangle(t) &= \begin{pmatrix}
				\psi(t)^\dagger J_x \psi(t)\\
				\psi(t)^\dagger J_y \psi(t)\\
				\psi(t)^\dagger J_z \psi(t)
			\end{pmatrix},
		\end{align}

		where \(\cdot^\dagger\) is the adjoint operator (also referred to as Hermitian conjugate, and is implemented as complex conjugate transpose). The expected spin projection can be interpreted as the average direction that the spin system is oriented in space when many systems of the equivalent state are measured (due to the Heisenberg uncertainty principle, only one component of the exact orientation of a quantum state can be known at any point in time, which is why we need to deal with expected values rather than absolute values). However, it can also be used to model the overall orientation of an ensemble of many quantum systems, like a BEC, for instance. \texttt{spinsim} has the functionality to calculate the expected spin projection of a system from its state.

	\subsection{Unitary time evolution and the Magnus expansion}
		Given that \(\psi(t)\) is a unit vector, it is possible to write \(\psi(t)\) in terms of a unitary transformation \(U(t, t_0)\) of the state \(\psi(t_0)\), for any time \(t_0\). In other words, for any \(t_0,t \in \mathbb{R}\), there is a unitary transformation \(U(t, t_0)\) such that

		\begin{align}
			\psi(t) &= U(t, t_0)\psi(t_0).
		\end{align}

		From Equation \eqref{eq:schroedinger}, \(U(t, t_0)\) also follows the Schroedinger equation via

		\begin{align}
			\frac{\mathrm{d}}{\mathrm{d}t}U(t, t_0)\psi(t_0) &= -i2\pi H(t)U(t, t_0)\psi(t_0)\textrm{ for any }\psi(t_0)\in\mathbb{C}^N,\textrm{ so}\\
			\frac{\mathrm{d}}{\mathrm{d}t}U(t, t_0) &= -i2\pi H(t)U(t, t_0).\label{eq:unitary_schroedinger}
		\end{align}

		Thus, to solve Equation \eqref{eq:schroedinger} for \(\psi(t)\) given a particular \(\psi(t_0)\), one only needs to solve Equation \eqref{eq:unitary_schroedinger} for \(U(t, t_0)\). Note that because \(U(t, t_0)\) is unitary, the calculated value of \(\psi(t)\) using this method will not change in magnitude, conserving probability.

		Also since \(U(t, t_0)\) is unitary (and continuous in time \(t\)), it can be written as a matrix exponential of a skew Hermitian matrix. For example, if the Hamiltonian \(H(t)\) is constant in time (and a real linear combination of Hermitian matrices defined by Equations \eqref{eq:spin_half_operators} or \eqref{eq:spin_one_operators}), then the solution would be

		\begin{align}
			U(t, t_0) &= \exp(-i2\pi H(t - t_0)).
		\end{align}

		The act of integrating a linear differential equation by exponentiation can be generalised for time dependent Hamiltonians using the Magnus Series \cite{blanes_magnus_2009}, where

		\begin{align}
			U(t, t_0) &= \exp(-i2\pi G(t - t_0))\\
			&= \exp(-i2\pi \sum_{m = 1}^\infty G_m(t - t_0)),
		\end{align}

		where \(G(t)\) is a series of \(G_m(t)\), each of which are a function of \(H\) and its commutators. \(G_1(t)\), for example, is just \(G_1(t) = H_\mathrm{av}t\), where \(H_\mathrm{av}\) is the average value of \(H\) over the interval being integrated over, which is the solution for when \(H\) is constant in time. The terms \(G_m(t)\) for \(m > 1\) are corrections to this approximation. One can truncate the Magnus series before infinity to make a Mangus expansion, which can given an approximation for \(U(t, t_0)\). Importantly, any truncation of the Magnus expansion is unitary, meaning that probability will be conserved when used.

		Something else to note about unitary time evolution operators, is that they can be broken into a product of multiple other time evolution operators via

		\begin{align}
			U(t, t_0) &= U(t, t_1)U(t_1, t_0).
		\end{align}

		Each of these parts can be approximated as a Magnus expansion. So, if the Magnus expansion of time evolution operator is inaccurate, it can be broken down into a product of many time evolution operators which each make a much smaller step between times. These components will each have a much more accurate Magnus expansion.

		In fact, many unitary time stepping techniques have been developed based on the Magnus expansion \cite{auer_magnus_2018}. Some of these techniques use Gauss-Legendre quadrature sampling to avoid the integration of commutators of the Hamiltonian, producing simple expressions that can be used for unitary time stepping. In particular, we used the commutator free 4 (CF4) from \cite{auer_magnus_2018} as our integration method. Suppose we wish to evaluate a short time step of \(U(t_f + \mathrm{d}t, t_f)\) by taking a short step \(\mathrm{d}t\) from a fine sample time \(t_f\). Then as part of the CF4 method, the field functions are sampled at particular times based on the second order Gauss-Legendre quadrature, given by

@@

		\begin{align}
			\overline{G}_1 &= 2\pi\mathrm{d}t\left(\frac{3 + 2\sqrt{3}}{12}H(t_1) + \frac{3 - 2\sqrt{3}}{12}H(t_2)\right)\textrm{, and}\label{eq:cf4_sample_1}\\
			\overline{G}_2 &= 2\pi\mathrm{d}t\left(\frac{3 - 2\sqrt{3}}{12}H(t_1) + \frac{3 + 2\sqrt{3}}{12}H(t_2)\right),\label{eq:cf4_sample_2}
		\end{align}

@@

		\begin{align}
			U(t_f + \mathrm{d}t, t_f) = \exp(-i\overline{G}_2)\exp(-i\overline{G}_1).\label{eq:cf4_implementation}
		\end{align}

	\subsection{Exponentiator}
		From Equation \eqref{eq:cf4_implementation}, evaluating matrix exponentials is a core part of the integration algorithm. However, the \texttt{spinsim} package uses the user provided field functions \(f(t)\) as a vector representation of the Hamiltonian of the spin system being simulated, rather than generic matrix elements in an array. This allows for implementation of matrix exponentiation that takes advantage of this representation. For all implementations of exponentiation, the \texttt{spinsim} computes the matrix exponential

		\begin{align}
			E(g) &= E(g_x, g_y, g_z, g_q)\\
			&= \exp(-i (g_x J_x + g_y J_y + g_z J_z + g_q Q)).
		\end{align}

		In practice, \(g\) is the Lie algebra (spin and quadrature) basis representation of \(\overline{G_1}\) or \(\overline{G_2}\) defined in Equations \eqref{eq:cf4_sample_1} and \eqref{eq:cf4_sample_2}, which need to be exponentiated in Equation \eqref{eq:cf4_implementation}. For spin half, the default exponentiator is in an analytic form. For spin one, an exponentiator based on the Lie Trotter product formula \cite{moler_nineteen_2003} is used instead. Explicitly, this formula is

		\begin{align}
			\exp\left( A + B\right) &= \lim_{n\to\infty} \left(\exp\left(\frac{A}{n}\right) \exp\left(\frac{B}{n}\right)\right)^n
		\end{align}

		Where \(A,B\in\mathbb{C}^{N\times N}\) are matrix operators, and \(n\in\mathbb{N}\). The equality holds for without the limit for all \(n\) if the operators commute, like it does for real and complex numbers. This tells us that we can approximate a matrix exponential of a linear combination of spin operators by reducing the size of the coefficients of these operators, taking the analytically simple exponentials of each of the operators \(-iJ_x, -iJ_y, -iJ_z\) and \(-iQ\) separately, multiplying each individual result together, and then raising the combination \(T\) to the power of the original reduction.

		For our spin one case, the exponential \(E(g)\) can be approximated as, for large \(2^\tau\),

>>>>

\section{Abstract}
	\emph{spinsim} is a \emph{python} package that simulates spin half and spin one quantum mechanical systems following a time dependent Shroedinger equation.
	It makes use of \texttt{numba.cuda} \cite{lam_numba_2015}, which is an \emph{LLVM} (Low Level Virtual Machine) \cite{lattner_llvm_2004} compiler, and other optimisations, to allow for fast and accurate evaluation on \emph{Nvidia Cuda} \cite{nickolls_scalable_2008} compatible systems using GPU parallelisation.
	\emph{spinsim} is available for installation on \emph{PyPI}, and the source code is available on \emph{github}.
	The initial use case for the package will be to simulate quantum sensing-based Bose Einstein condensate (BEC) experiments for the Monash University School of Physics and Astronomy spinor BEC lab, but we anticipate it will be useful in simulating any range of spin half or spin one quantum systems with time dependent Hamiltonians that cannot be solved analytically.
	These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin one systems of nitrogen vacancy centres (NVCs) and BECs.

@@

\section{Implementation and architecture}
	% \subsection{Mathematical methods}
	\subsection{Quantum mechanics background}
		\emph{spinsim} solves the time-dependent Schr\"{o}dinger equation
		\begin{align}
			i\hbar\frac{\mathrm{d}\psi(t)}{\mathrm{d}t} &= H(t)\psi(t),\label{eq:schroedinger}
		\end{align}
		where the quantum state \(\psi(t) \in \mathbb{C}^N\) is assumed normalised, and the Hamiltonian \(H(t) \in \mathbb{C}^{N \times N}\) is Hermitian.
		Here \(N\) is the number of levels in the quantum system.
		Often we are considering systems with spin \(J\) of half or one.
		Here spin-half is the \(N = 2\) case, and spin-one is \(N = 3\).
		We set \(\hbar = 1\), so that the Hamiltonian has physical dimension of inverse frequency.
		% Furthermore, we have chosen for the Hamiltonian \(H(t)\) to have the physical dimension of frequency, rather than the more typical dimension of energy. To elaborate, we represent energies \(E\) of the system as their equivalent frequencies \(f\) from the Planck relation

		% \begin{align}
		% 	E &= \hbar \omega\\
		% 	&= 2\pi\hbar f,
		% \end{align}

		% where \(\hbar\) is (reduced) Planck's constant (or equivalently, setting 
		% \(\hbar = 1\)). In addition, rather than being represented by standard coordinates in \(\mathbb{C}^{N \times N}\), in \emph{spinsim} the Hamiltonian \(H(t)\) is instead represented with respect to a choice of basis for the corresponding Lie Algebra, \(\mathfrak{su}(N)\). For example, when set to spin half mode, the \emph{spinsim} package solves the time dependent Schroedinger equation of the form

		Rather than parametrizing the problem in terms of the matrix elements of \(H(t)\), we consider time varying real coefficients in a linear combination of fixed operators,
		\begin{align}
			H(t) &= \sum_{j = 1}^{N^2 - 1} \omega_j(t) O_j.
		\end{align}
		It is well known that a charged spin half system with magnetic moment \(\overrightarrow{\mu}\), and gyromagnetic ratio \(\gamma\), in a magnetic field \(\overrightarrow{B}(t)\), has Hamiltonian
		\begin{align}
			H(t) &= -\overrightarrow{B}(t)\cdot \overrightarrow{\mu}\\
			&= -\gamma \overrightarrow{B}(t)\cdot \overrightarrow{J}\\
			&= -\gamma \left(B_x(t) J_x + B_y(t) J_y + B_z(t) J_z\right)\\
			&= \left(-\gamma B_x(t) J_x\right) + \left(-\gamma B_y(t) J_y\right) + \left(-\gamma B_z(t) J_z\right)\\
			&= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z.
		\end{align}

		Here \(J_x, J_y\) and \(J_z\) are spin operators, equal to the Pauli matrices scaled down by \(\frac12\).
		We exclude the identity so that the Hamiltonian is traceless, which we can do because it corresponds to choosing an energy zero point, which is physically meaningless.

		In the spin one case there is no standard basis of operators \(A_j\).
		Choices include the Gell Mann matrices, and multiple dipole-quadrupole bases \cite{hamley_spin-nematic_2012}\cite{di_dipolequadrupole_2010}.
		In general, we can choose any basis from the 8-dimensional Lie algebra \(\mathfrak{su}(3)\), which is the vector space of traceless Hermitian operators that can generate transformations (from the corresponding Lie group \(SU(3)\)) in the spin one system.
		With this in mind, we choose to represent the Hamiltonian a linear combination of matrices from a 4-dimensional subspace of \(\mathfrak{su}(3)\), consisting of the spin matrices \(J_x, J_y\) and \(J_z\), and a single quadrupole operator \(Q = \frac13\mathrm{diag}(1, -2, 1)\),
		\begin{align}
			H(t) &= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q.
		\end{align}

		Note that \(Q\) is proportional to \(Q_{zz}\) \cite{hamley_spin-nematic_2012} and \(Q_0\) \cite{di_dipolequadrupole_2010} from alternative quadrupole bases.
		These are the only operators necessary to simulate many spin one quantum systems, including quadratic Zeeman splitting described by the Breit-Rabi formula.
		\emph{spinsim} can also be configured to solve a general spin one system by setting the Hamiltonian to an arbitrary point in \(\mathfrak{su}(3)\), using the full quadrupole basis, which extends the possible Hamiltonian to
		\begin{align}
			H(t) &= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q\nonumber\\
			&+ \omega_{ux}(t) U_1 + \omega_{uy}(t) U_2 + \omega_{vx}(t) V_1 + \omega_{vy}(t) V_2.
		\end{align}
		Here the added operators are those defined in \cite{di_dipolequadrupole_2010}. Note that this is included for completeness, but has not been thoroughly tested, as our lab has no physical context for modelling the \(U_1, U_2, V_1\) and \(V_2\) operators.
		% In general ... Lie algebra ... mathfrak ... admits a basis ... .
		% \begin{align}
		% 	\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z) \psi(t),
		% \end{align}

		% where \(i^2 = -1\), \(\psi(t) \in \mathbb{C}^2\), and the spin half spin projection operators are given by

		% DONT WRITE OUT, sigma
		% \begin{align}
		% 	J_x &= \frac12\begin{pmatrix}
		% 		0 & 1 \\
		% 		1 & 0
		% 	\end{pmatrix},
		% 	&J_y &= \frac12\begin{pmatrix}
		% 		0 & -i \\
		% 		i &  0
		% 	\end{pmatrix},
		% 	&\textrm{and }J_z &= \frac12\begin{pmatrix}
		% 		1 &  0 \\
		% 		0 & -1
		% 	\end{pmatrix}.\label{eq:spin_half_operators}
		% \end{align}

		% Call omega, no implementation detail
		% The fields \(f\) that represents the time dependent Hamiltonian \(H(t)\), are the collection of energy functions \(f_x(t), f_y(t), f_z(t)\), with time \(t\) and \(f\) with physical dimension of frequency that control the dynamics of the system. The user must define a method that returns a sample of these field functions when a sampling time is input. In physical terms, these functions could represent the \(x,y,z\) components of a magnetic field applied to a magnetically spin one BEC or NVC.

		% Similarly, when \emph{spinsim} is set to spin one mode, it can solve the Schroedinger equation of the form

		% \begin{align}
		% 	\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z + f_q(t) Q) \psi(t).
		% \end{align}

		% where now \(\psi(t) \in \mathbb{C}^3\), and the spin one operators are given by

		% \begin{align}
		% 	J_x &= \frac{1}{\sqrt{2}}\begin{pmatrix}
		% 		0 & 1 & 0 \\
		% 		1 & 0 & 1 \\
		% 		0 & 1 & 0
		% 	\end{pmatrix},&
		% 	J_y &= \frac{1}{\sqrt{2}}\begin{pmatrix}
		% 		0 & -i &  0 \\
		% 		i &  0 & -i \\
		% 		0 &  i &  0
		% 	\end{pmatrix},\nonumber\\
		% 	J_z &= \begin{pmatrix}
		% 		1 & 0 &  0 \\
		% 		0 & 0 &  0 \\
		% 		0 & 0 & -1
		% 	\end{pmatrix},&
		% 	\textrm{and }Q &= \frac{1}{3}\begin{pmatrix}
		% 		1 &  0 & 0 \\
		% 		0 & -2 & 0 \\
		% 		0 &  0 & 1
		% 	\end{pmatrix}.\label{eq:spin_one_operators}
		% \end{align}

		% The matrices \(J_x, J_y, J_z\) are regular spin operators, and \(Q\) is a quadrupole operator. Note that \(Q\) is proportional to \(Q_{zz}\) as defined by Hamley et al \cite{hamley_spin-nematic_2012}, and \(Q_0\) as defined by Di et al \cite{di_dipolequadrupole_2010}.
		% INLINE Q DIAG. WHY WE'VE PICKED THESE FOUR. For ground state spin one atoms in a magnetic field, ... Breit Rabi ... . COULD BE READILY EXTENDED TO INCLUDE FURTHER ELEMENTS OR TO HIGHER SPINS.

		% Too far out of scope
		% Alternative: integrate Bloch equations 3 coupled real equations
		% fast and parallel function from wavefunctions
		As well as integrating the Schr\"odinger equation, \emph{spinsim} also has the functionality to calculate the expected spin projection \(\left\langle \overrightarrow{J}\right\rangle\)(t) of a system from its state.
		In an experimental setting, one cannot measure the value of the state itself, and must instead measure observables such as spin projection.
		If this is done for an ensemble of systems, the average spin projection over all systems will converge to the expected value as calculated here.

		% Frequently when one calculates the times series of the quantum state \(\psi(t)\), they are also interested in its expected spin projection. It is given by the vector

		% % Inline
		% \begin{align}
		% 	\left\langle J\right\rangle(t) &= \begin{pmatrix}
		% 		\psi(t)^\dagger J_x \psi(t)\\
		% 		\psi(t)^\dagger J_y \psi(t)\\
		% 		\psi(t)^\dagger J_z \psi(t)
		% 	\end{pmatrix},
		% \end{align}

		% where \(\cdot^\dagger\) is the adjoint operator. The expected spin projection can be interpreted as the average direction that the spin system is oriented in space when many systems of the equivalent state are measured (due to the Heisenberg uncertainty principle, only one component of the exact orientation of a quantum state can be known at any point in time, which is why we need to deal with expected values rather than absolute values). However, it can also be used to model the overall orientation of an ensemble of many quantum systems, like a BEC, for instance. \emph{spinsim} has the functionality to calculate the expected spin projection of a system from its state.

	\subsection{Unitary time evolution and the Magnus expansion}
		It is possible to write \(\psi(t)\) in terms of a unitary transformation \(\mathcal{U}(t, t_0)\) of the state \(\psi(t_0)\), for any time \(t_0\), that is, \(\psi(t) = \mathcal{U}(t, t_0)\psi(t_0)\).
		It then follows from Equation \eqref{eq:schroedinger} that \(\mathcal{U}(t, t_0)\) also follows a Schr\"{o}dinger equation,
		\begin{align}
			i\hbar\frac{\mathrm{d}\mathcal{U}(t, t_0)}{\mathrm{d}t} &= H(t)\mathcal{U}(t, t_0).\label{eq:unitary_schroedinger}
		\end{align}

		Thus, to solve Equation \eqref{eq:schroedinger} for \(\psi(t)\) given a particular \(\psi(t_0)\), one only needs to solve Equation \eqref{eq:unitary_schroedinger} for \(\mathcal{U}(t, t_0)\).
		% Note that because \(\mathcal{U}(t, t_0)\) is unitary, the calculated value of \(\psi(t)\) using this method will not change in magnitude, conserving probability.
		Also since \(\mathcal{U}(t, t_0)\) is unitary, it can be written as a matrix exponential of a skew-Hermitian (also termed anti-Hermitian) matrix.
		If the Hamiltonian is constant, then this exponential is % Check if continuity requirement on unitary
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp(-i H\cdot(t - t_0)).\label{eq:exp_sol_of_constant}
		\end{align}
		% The act of integrating a linear differential equation by exponentiation can be generalised for time dependent Hamiltonians using the Magnus Series \cite{blanes_magnus_2009}, where
		For a time varying Hamiltonian, the general solution for \(\mathcal{U}(t, t_0)\) is much more complex, because it encapsulates the full solution to the time-dependent Schr\"{o}dinger equation.
		The Dyson series gives an explicit expression for \(\mathcal{U}(t, t_0)\) in terms of multiple time integrals over nested time commutators of \(H(t')\). % Cite 2? recent papers
		While the Dyson series has been used numerically, it has only been so recently, in particular because once truncated, it is in general no longer unitary.

		The Magnus series
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp\left(\Omega(t, t_0)\right) = \exp\left(\sum_{m = 1}^\infty \Omega_m(t, t_0)\right)
		\end{align}
		explicitly preserves unitarity when truncated\cite{magnus_exponential_1954}.
		The first terms in the Magnus series are\cite{blanes_magnus_2009}
		\begin{align}
			\Omega_1(t, t_0) &= \int_0^t\mathrm{d}t_1A(t_1)\\
			\Omega_2(t, t_0) &= \frac12\int_{t_0}^t\mathrm{d}t_1\int_{t_0}^{t_1}\mathrm{d}t_2[A(t_1), A(t_2)]\\
			\Omega_3(t, t_0) &= \frac16\int_{t_0}^t\mathrm{d}t_1\int_{t_0}^{t_1}\mathrm{d}t_2\int_{t_0}^{t_2}\mathrm{d}t_3\left([A(t_1), [A(t_2), A(t_3)]] + [A(t_3), [A(t_2), A(t_1)]]\right),
		\end{align}
		where\([X, Y] = XY - YX\) is the commutator, and, in the case of quantum mechanics, \(A(t') = -iH(t')\). Note that, when truncated to first order, the magnus expansion \(\mathcal{U}(t, t_0) = \exp\left(\int_0^t\mathrm{d}t_1\left(-iH(t_1)\right)\right)\) reduces to Equation \eqref{eq:exp_sol_of_constant}, with \(H\) approximated to be constant with its average value.

		% Insert first few omegas here
		% Throw back to constant
		% Here each \(\Omega_m(t)\) are integrals over time commutators of \(H(t')\).
		% \(\Omega_1(t)\), for example, is \(\Omega_1(t) = H_\mathrm{av}t\), where \(H_\mathrm{av}\) is the average value of \(H\) over the interval being integrated over, which is the solution for when \(H\) is constant in time.
		% The terms \(\Omega_m(t)\) for \(m > 1\) are corrections to this approximation. One can truncate the Magnus series before infinity to make a Mangus expansion, which can given an approximation for \(\mathcal{U}(t, t_0)\).
		% Importantly, any truncation of the Magnus expansion is unitary, meaning that probability will be conserved when used.

		A Magnus series over a time step \(T\) will be guaranteed to converge only if the spectral norm of \(H\) is sufficiently small, \(\int_{t_0}^{t_0 + T}\mathrm{d}t_1\|H(t_1)\| < \pi\)\cite{blanes_magnus_2009}. Furthermore, each subsequent term in the expansion increases in complexity rapidly. For these reasons, the Magnus expansion can be used as a time stepping method rather than a single step to solve the complete system. Time stepping can be made from the fact that time evolution operators can be split into a product via
		\begin{align}
			\mathcal{U}(t, t_0) &= \mathcal{U}(t, t_{n - 1})\mathcal{U}(t_{n - 1}, t_{n - 2})\cdots\mathcal{U}(t_2, t_1)\mathcal{U}(t_1, t_0).\label{eq:product_of_time_evolution}
		\end{align}
		Each of the time evolution operators can be approximated by a Magnus expansion.

		% Each of these parts can be approximated as a Magnus expansion. So, if the Magnus expansion of time evolution operator is inaccurate, it can be broken down into a product of many time evolution operators which each make a much smaller step between times. These components will each have a much more accurate Magnus expansion. % convergence of Magnus series, not simple. Also Dyson. Standard bounds.
		% Numerical solver takes only a few terms of the expansion. However, any long operator can be broken into a product. Cite Magnus papers. Designing a Magnus based integrator is a trade off.

		Many unitary time stepping techniques have been developed based on the Magnus expansion \cite{auer_magnus_2018}.
		Some of these techniques use Gauss-Legendre quadrature sampling to avoid the integration of commutators of the Hamiltonian, producing simple expressions that can be used for unitary time stepping.
		In particular, we used the commutator free 4 (CF4) from \cite{auer_magnus_2018} as our integration method.
		Suppose we wish to evaluate a short time step of \(\mathcal{U}(t_f + \mathrm{d}t, t_f)\) by taking a short step \(\mathrm{d}t\) from a fine sample time \(t_f\).
		Then as part of the CF4 method, the field functions are sampled at particular times based on the second order Gauss-Legendre quadrature, given by

@@

		\begin{align}
			\overline{H}_1 &= \mathrm{d}t\left(\frac{3 + 2\sqrt{3}}{12}H(t_1) + \frac{3 - 2\sqrt{3}}{12}H(t_2)\right)\textrm{, and}\label{eq:cf4_sample_1}\\
			\overline{H}_2 &= \mathrm{d}t\left(\frac{3 - 2\sqrt{3}}{12}H(t_1) + \frac{3 + 2\sqrt{3}}{12}H(t_2)\right),\label{eq:cf4_sample_2}
		\end{align}

@@

		\begin{align}
			\mathcal{U}(t_f + \mathrm{d}t, t_f) = \exp(-i\overline{H}_2)\exp(-i\overline{H}_1).\label{eq:cf4_implementation}
		\end{align}

	\subsection{Exponentiator}
		From Equation \eqref{eq:cf4_implementation}, evaluating matrix exponentials is a core part of the integration algorithm.
		However, the \emph{spinsim} package uses the user provided field functions \(\omega(t)\) as a vector representation of the Hamiltonian of the spin system being simulated, rather than generic matrix elements in an array.
		This allows for implementation of matrix exponentiation that takes advantage of this representation.
		For all implementations of exponentiation, the \emph{spinsim} computes the matrix exponential

		\begin{align}
			E(\omega) &= E(\omega_x, \omega_y, \omega_z, \omega_q)\\
			&= \exp(-i (\omega_x J_x + \omega_y J_y + \omega_z J_z + \omega_q Q)).
		\end{align}

		In practice, \(\omega\) is the Lie algebra (spin and quadrature) basis representation of \(\overline{H_1}\) or \(\overline{H_2}\) defined in Equations \eqref{eq:cf4_sample_1} and \eqref{eq:cf4_sample_2}, which need to be exponentiated in Equation \eqref{eq:cf4_implementation}.
		For spin half, the default exponentiator is in an analytic form.
		For spin one, an exponentiator based on the Lie Trotter product formula \cite{moler_nineteen_2003} is used instead.
		Explicitly, this formula is

		\begin{align}
			\exp\left( X + Y\right) &= \lim_{n\to\infty} \left(\exp\left(\frac{X}{n}\right) \exp\left(\frac{Y}{n}\right)\right)^n
		\end{align}

		Where \(X,Y\in\mathbb{C}^{N\times N}\) are matrix operators, and \(n\in\mathbb{N}\).
		The equality holds for without the limit for all \(n\) if the operators commute, like it does for real and complex numbers.
		This tells us that we can approximate a matrix exponential of a linear combination of spin operators by reducing the size of the coefficients of these operators, taking the analytically simple exponentials of each of the operators \(-iJ_x, -iJ_y, -iJ_z\) and \(-iQ\) separately, multiplying each individual result together, and then raising the combination \(T\) to the power of the original reduction.
